\texttt Design date

### 2017-03-16 latex/algorithmdesign/cipherconstructions.tex
SIMD testing

\section{First section}

Meet Bob and Alice. Bob and Alice are pen pals; they write letters to each other, and send them through the mail. Additionally, we will introduce Eve:  Eve has no respect for others, as well as a pathological need to try and do wrong; She is not opposed to lying, cheating, and stealing, and is very creative in this arts. She will go to great length to do these things if she thinks it will benefit her; Even when there is no real benefit to her, she will do such things whenever she can get away with it, simply because she can.

Unfortunately, Eve is the mail carrier. Fortunately, Bob and Alice are aware of Eve's nature, and they seek to find ways of communicating with each other without Eve interfering. In order to do so, they need to enumerate all of the ways that Eve could abuse their communications, and develop appropriate counter-measures.

\subsection{Lying, Cheating, and Stealing}

The first and most obvious concern is Eve opening and reading the messages that are exchanged between Bob and Alice.

Second, and slightly more devious, is that Eve could manipulate messages sent between the two.

Third, Eve actually does not have to wait for Bob to actually send a letter to Alice. She can simply write her own letter, put Bob's name on it, and deliver it to Alice.

Fourth, she could withhold or make copies of messages, and save them for delivery at some other arbitrary point in time.

Finally, Eve could simply choose not to deliver the messages at all.

\subsection{Solutions}

For example, Bob might send Alice a christmas card with money inside, and a note that says "I am sending you \$20; I hope you have a nice holiday". Upon opening this, Eve could simply take the money, and leave a new note in the card that say "Sorry, I can't afford gifts this year; I hope you have a nice holiday anyways. PS: Please send me some money". Then she could easily repeat this process.

>>>>

\section{Cipher Constructions}

The design of a cipher can be split into (at least) two parts: The cipher construction, and the core permutation. The two archetypal examples of cipher construction are the Feistel Network and Substiution-Permutation Network.

In the former, the state is divided into two sections that are typically of equal size; A key is added to some side(s), then a core "randomizing" function is applied to one half to produce a key stream; This key stream is then combined with the other half, in order to encipher it. We can see then, that the core "randomizing function" is interchangeable within the Feistel Network construction.

In a Substitution-Permutation network, the state is typically subdivided into many smaller sections. The algorithm typically consists of addition of a key layer, followed by at least one application of an invertible "randomizing" function applied to the entire state; This is repeated until a sufficient "security margin" is obtained. This construction can be modeled as the composition of two functions: The key addition function, and the application of the core permutation on the state. As in the Feistel Network, the core permutation is interchangeable within the Substitution-Permutation network construction.

We can see then, that there are at least two, effectively separate aspects to cipher design: The cipher construction, and the core permutation. This work focuses on the former. While the core permutation is critical for resistance to generic attacks such as linear and differential cryptanalysis, there also exist generic attacks against the cipher constructions. Additionally, while the design of the permutation can be hugely influential on the performance of the algorithm, so too can the cipher construction.

\section{Construction Design Considerations}

First, the target platform needs to be established. What is ideal for constrained hardware is not necessarily ideal for a modern CPU. In addition to the target platform(s), the performance goals of the algorithm should be outlined: Is latency important, or is throughput for large messages important? Is it important to use as little (or much) RAM?

Less obvious concerns that are still very important include: Is resistance to side channel leakage minimized? Does it facilitate or impede attempts to implement counter-measures?

"The simple" - RC4 like (shuffling huge array of words)

"The small" - 8-bit friendly/constrained device friendly.

"The fast" - salsa like (SIMD with big registers)

\subsection{A subsection}

More text.

### 2017-03-16 latex/algorithmdesign/untitled-4.tex
SIMD testing

Unfortunately, Eve is the mail carrier. Fortunately, Bob and Alice are aware of Eve's nature, and they seek to find ways of communicating with each other without Eve interfering. In order to do so, they need to enumerate all of the ways that Eve could abuse their communications, and develop appropriate counter-measures.

@@

For example, Bob might send Alice a christmas card with money inside, and a note that says "I am sending you \$20; I hope you have a nice holiday". Upon opening this, Eve could simply take the money, and leave a new note in the card that say "Sorry, I can't afford gifts this year; I hope you have a nice holiday anyways. PS: Please send me some money". Then she could easily repeat this process.

>>>>

Unfortunately, Eve is the mail carrier, and is responsible for delivering messages to/from the pair. Fortunately, Bob and Alice are aware of Eve's nature, and they seek to find ways of communicating with each other without Eve interfering. In order to do so, they need to enumerate all of the ways that Eve could abuse their communications, and develop appropriate counter-measures.

@@

In order to preclude Eve's ability to read their messages, Bob and Alice realize they need some way to obscure the information such that only the intended recipient may recover the actual information from the transported information. They need some kind of set of steps that they can apply to their message that will: 1: Produce output that is meaningless, and 2: The steps can be undone by the intended receipient to recover the actual message.

So clearly, they reason, they need some kind of technique that will allow the two of them be able to perform the steps, but nobody else (i.e. Eve) should be able to perform the steps. After some thought, they realize that if they can keep some part of the steps secret, and only the two of them know the secret steps, then clearly Eve will not be able perform the transformation and recover the message. The pair realize that Eve will go out of her way to attempt to figure out the secret steps, and that they will likely need to include many secrets; If they only use one secret, Eve could easily guess it.

% communicating the coin flip

### 2017-03-16 latex/cipher_templates/blockcipher/arx/template.tex
SIMD testing

>>>>

The ARX construction is one method for designing a block cipher. ARX based ciphers forego the use of lookup tables as a source of non-linearity, utilizing integer addition and boolean functions combined with rotations and xor to obtain the required source of "confusion". ARX designs are generally simple to implement without exposing timing based side channels. They tend to be efficient in software on a modern CPU, as many designs can utilize the parallelism offered by large registers and vectorized instructions.
